A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-10-19 doc/large_strain.tex
improve MF algorithms

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$, current FE solution $\overline{\gz u^h}$,
  cached $c_1 := \mu - 2 \lambda \log(J)$ for each cell and quadrature point}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $\gz F = \gz I + \gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $J = \rm{det}(\gz F)$ \tcp*{scalar}
          evaluate $\gz b = \gz F \cdot \gz F^T$ \tcp*{2nd order symmetric}
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\gz \tau = \mu \gz b - c_1 \gz I$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{G}} \gz g = \gz g \cdot \gz \tau/J$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}}$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s = \left[2 c_1 \gz g_s + 2 \lambda\, \rm{tr}( \gz g_s) \gz I \right]/J$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ for contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
      evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator: scalar cache}
  \label{alg:mf_scalar}
\end{algorithm}

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached
  $c_1 := 2\left[\mu - 2 \lambda \log(J)\right]/J$,
  $c_2 := 2\lambda/J$ and $\gz \tau/J$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\boldsymbol{\mathcal{G}}\gz g = \gz g \cdot \left[\gz \tau/J \right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$ \;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}} \gz g_s = c_1 \gz g_s + c_2 \, \rm{tr}(\gz g_s) \gz I$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}} \gz g_s $ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator, variant 1: cache second-order Kirchhoff stress $\gz \tau$ and thereby avoid the need to evaluate refential quantities like $\gz F$ at runtime.}
  \label{alg:mf_tensor2}
\end{algorithm}

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached $\gz \tau/J$ and $\boldsymbol{\mathcal{C}}$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
          evaluate $\boldsymbol{\mathcal{G}} \gz g = \gz g \cdot \left[\gz \tau/J\right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s = \boldsymbol{\mathcal{C}}\gz : \gz g_s$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}}\gz g_s$ for contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator: cache fourth-order material tangent $\boldsymbol{\mathcal{C}}$ and Kirchhoff stress $\gz \tau$.}
  \label{alg:mf_tensor4}
\end{algorithm}

>>>>

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$, current FE solution $\overline{\gz u^h}$,
  cached $c_1 := \mu - 2 \lambda \log(J)$ for each cell and quadrature point}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        evaluate the following tensors at each quadrature point using sum factorization \;
        $\gz \nabla_X \gz u^h$ \tcp*{2nd order}
        $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
        $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\gz F = \gz I + \gz \nabla_X \gz u^h$ \tcp*{2nd order}
          evaluate $J = \rm{det}(\gz F)$ \tcp*{scalar}
          evaluate $\gz b = \gz F \cdot \gz F^T$ \tcp*{2nd order symmetric}
          evaluate $\gz \tau = \mu \gz b - c_1 \gz I$ \tcp*{2nd order symmetric}
          evaluate $\boldsymbol{\mathcal{G}} \gz g := \gz g \cdot \gz \tau/J$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s := \left[2 c_1 \gz g_s + 2 \lambda\, \rm{tr}( \gz g_s) \gz I \right]/J$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ for contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
      evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator: scalar cache}
  \label{alg:mf_scalar}
\end{algorithm}

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached
  $c_1 := 2\left[\mu - 2 \lambda \log(J)\right]/J$,
  $c_2 := 2\lambda/J$ and $\gz \tau/J$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        evaluate the following tensors at each quadrature point using sum factorization \;
        $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
        $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\boldsymbol{\mathcal{G}}\gz g := \gz g \cdot \left[\gz \tau/J \right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$ \;
          evaluate $\boldsymbol{\mathcal{C}} \gz g_s := c_1 \gz g_s + c_2 \, \rm{tr}(\gz g_s) \gz I$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}} \gz g_s $ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator, variant 1: cache second-order Kirchhoff stress $\gz \tau$ and thereby avoid the need to evaluate refential quantities like $\gz F$ at runtime.}
  \label{alg:mf_tensor2}
\end{algorithm}

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached $\gz \tau/J$ and $\boldsymbol{\mathcal{C}}$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        evaluate the following tensors at each quadrature point using sum factorization \;
        $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
        $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\boldsymbol{\mathcal{G}} \gz g := \gz g \cdot \left[\gz \tau/J\right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$\;
          evaluate $\boldsymbol{\mathcal{C}}\gz g_s := \boldsymbol{\mathcal{C}}\gz : \gz g_s$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}}\gz g_s$ for contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator: cache fourth-order material tangent $\boldsymbol{\mathcal{C}}$ and Kirchhoff stress $\gz \tau$.}
  \label{alg:mf_tensor4}
\end{algorithm}
